---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bddR

<!-- badges: start -->
<!-- badges: end -->

The goal of `bddR` is to provide [Gherkin syntax](https://cucumber.io/docs/gherkin/) helpers for [`testthat`'s `describe()` and `it()` functions](https://testthat.r-lib.org/reference/describe.html).

## Installation

You can install the development version of bddR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mjfrigaard/bddR")
```

```{r}
library(bddR)
```


```{r testthat, eval=TRUE}
if (!require(pak)) {
  install.packages("pak",
    repos = "http://cran.us.r-project.org"
  )
}
pak::pkg_install("r-lib/testthat", upgrade = TRUE, ask = FALSE)
library(testthat)
```

## Example

`testthat`'s BDD functions allow for more explicit descriptions of tests: 

```{r}
describe(description = "verify that you implement the right things", code = {
  it(description = "ensure you do the things right", code = {
    expect_true(TRUE)
  })
})
```

`bddR` provides a set of helpers for writing [Gherkin-style](https://cucumber.io/docs/gherkin/reference/) features, backgrounds and scenarios which can be placed in `describe()` and `it()`:  

```{r}
describe(
  feature(
    title = "My function's feature",
    as_a = "user of this function",
    i_want = "to verify that I implemented the right thing",
    so_that = "to ensure I did the thing right"
  ), code = {
  it(
    scenario(
        title = "Example of thing",
        given = "My package is installed and loaded",
        when = "I do something",
        then = "the right thing happens"
      ), code = {
    expect_true(TRUE)
  })
})
```

`bddR` functions can also it can be dropped into the `desc` argument of `test_that()`:

```{r}
test_that(
  scenario(
        title = "Example of thing",
        given = "My package is installed and loaded",
        when = "I do something",
        then = "the right thing happens"
      ),
  code = {
    expect_true(TRUE)
  }
)
```

Check out the [Getting Started vignette](https://mjfrigaard.github.io/bddR/articles/start.html) to learn more.
