---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bddR

<!-- badges: start -->
<!-- badges: end -->

The goal of `bddR` is to provide [Gherkin syntax](https://cucumber.io/docs/gherkin/) helpers for [`testthat`'s `describe()` and `it()` functions](https://testthat.r-lib.org/reference/describe.html).

## Installation

You can install the development version of bddR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mjfrigaard/bddR")
```

```{r}
library(bddR)
```


```{r testthat, eval=FALSE}
if (!require(pak)) {
   install.packages("pak", 
     repos = "http://cran.us.r-project.org")
}
pak::pak("testthat")
library(testthat)
```

## Feature 

Adding [Gherkin keywords](https://cucumber.io/docs/gherkin/reference/) to `describe()` and `it()` can be tiresome. 

```{r}
testthat::test_that("
  Feature:
	  As a user
	  I want to see the changes in the plot
	  So that I can visualize the impact of my customizations
  ", code = {
  testthat::expect_true(TRUE)
})
```

`feature()` allows you to simply fill in the arguments: 

```{r}
feature(
    As_a = "user",
    I_want = "to see the changes in the plot", 
   So_that = "I can visualize the impact of my customizations")
```

It returns a `glue`/`character` string:

```{r}
class(feature(
    As_a = "user",
    I_want = "to see the changes in the plot", 
   So_that = "I can visualize the impact of my customizations"))
```

So it can be dropped into a test `description`:

```{r}
testthat::test_that(
  feature(
    As_a = "user",
    I_want = "to see the changes in the plot", 
   So_that = "I can visualize the impact of my customizations"), code = {
  testthat::expect_true(TRUE)
})
```

## Scenario 

Scenarios have `Given`, `When`, `Then` (and sometimes `And`) statements:

```{r}
testthat::describe(
  feature(
    As_a = "user",
    I_want = "to see the changes in the plot", 
   So_that = "I can visualize the impact of my customizations"), 
  code = {
    testthat::it("
      Scenario: 
          Given 
          When 
          Then 
      ", code = {
      testthat::expect_true(TRUE)
    })
  })
```



```{r}
testthat::describe(
    feature(As_a = "user",
            I_want = "to see the changes in the plot", 
            So_that = "I can visualize the impact of my customizations"),
  code = {
    testthat::it(
      scenario(Given = "I have launched the application",
         When = "I interact with the sidebar controls",
         Then = "the graph should update with the selected options"),
      code = {
      testthat::expect_true(TRUE)
    })
  })
```

We can add an `And` statement for more background/context:

```{r}
testthat::describe(
    feature(As_a = "user",
            I_want = "to see the changes in the plot", 
            So_that = "I can visualize the impact of my customizations"),
  code = {
    testthat::it(
      scenario(Given = "I have launched the application",
               And = "it contains movie review data from IMDB and Rotten Tomatoes",
               When = "I interact with the sidebar controls",
               Then = "the graph should update with the selected options"),
      code = {
      testthat::expect_true(TRUE)
    })
  })
```

More than one `And`? Just drop them in a `list()`:

```{r}
testthat::describe(
  feature(As_a = "user",
            I_want = "to see the changes in the plot", 
            So_that = "I can visualize the impact of my customizations"), 
  code = {
    testthat::it(
      scenario(Given = "I have launched the application",
               And = list("it contains movie review data from IMDB and Rotten Tomatoes",
                          "the data contains variables like 'Critics Score' and 'MPAA'"),
               When = "I interact with the sidebar controls",
               Then = "the graph should update with the selected options"),
      code = {
      testthat::expect_true(TRUE)
    })
  })
```

