---
title: "Bundling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bundling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(testthat)
```

```{r setup}
library(pickler)
```

Make sure you have the latest version of `testthat` installed.

```{r testthat, eval=FALSE, message=FALSE, warning=FALSE, results='hide'}
if (!require(pak)) {
  install.packages("pak",
    repos = "http://cran.us.r-project.org"
  )
}
pak::pkg_install("r-lib/testthat", upgrade = TRUE, ask = FALSE)
library(testthat)
```


## Bundling 

Ideally, `pickler`'s functions are developed independently and then wrapped in a `bundle()` before passing to the `testthat` BDD functions. Consider the `feature()` and `scenario()` below for the `split_cols()` function:

```{r , eval=FALSE}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
)
```

Bundling makes it easier to combine features and backgrounds into a single call:

```{r , echo=FALSE, eval=TRUE, comment=""}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
)
```

`bundle()` also allows us to pass a `feature()` and `background()` to `describe()`, then test a particular `scenario()` with `it()`:

```{r}
describe(
  # bundle feature and background
  bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
), code = {
    it(
    # test scenario 1
     scenario(
        title = "Split column with a default pattern",
        given = "a dataframe [data] with a specified column [col]",
        when = "I split the [col] column using the default [[^[:alnum:]]+]",
        then = "the [col] column should be split into multiple columns"
      ), code = {
      
      expect_true(TRUE)
      
    })
    it(
    # test scenario 2
     scenario(
        title = "Split column with a default pattern",
        given = "a dataframe [data] with a specified column [col]",
        when = "I split the [col] column using the default [[^[:alnum:]]+]",
        then = "the new columns should be named with the default prefix [cols_]"
      ), code = {
      
      expect_true(TRUE)
      
    })
  })
```

