---
title: "Getting started"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
```

This vignette covers how to use `pickler` functions with [`testthat`'s `describe()` and `it()` functions](https://testthat.r-lib.org/reference/describe.html).

```{r setup}
library(pickler)
```

Make sure you have the latest version of `testthat` installed.

```{r testthat, eval=TRUE, message=FALSE, warning=FALSE, results='hide'}
if (!require(pak)) {
  install.packages("pak",
    repos = "http://cran.us.r-project.org"
  )
}
pak::pkg_install("r-lib/testthat", upgrade = TRUE, ask = FALSE)
library(testthat)
```

## Motivation

`pickler` was motivated by the latest recommendation from the `testthat` authors on moving the test scope to *within* the test (i.e., remove or reduce the need for code outside the call to `test_that()`).

<br>

> ***To avoid code outside of `test_that()`:***
> 
> - ***Move file-scope logic to either narrower scope (just this test) or a broader scope (all files)***
> 
> - ***It's ok to copy and paste: test code doesn't have to be super dry. Obvious >> DRY***  - [Package Development Masterclass, posit::conf(2023)](https://github.com/posit-conf-2023/pkg-dev-masterclass/blob/main/materials/3-advanced-testing.pdf)

<br>

Adopting test [fixtures, helpers, and setup](https://www.tidyverse.org/blog/2020/04/self-cleaning-test-fixtures/) files will address the first item in the list above, but `pickler` is mostly focused on the second item: making tests obvious and clear. 

## Behavior-driven development functions

The behavior-driven development (BDD) [functions](https://testthat.r-lib.org/reference/describe.html) are excellent additions to the suite of testing tools provided by `testthat`. They work like so: 

```{r}
describe(description = "verify that you implement the right things", code = {
  it(description = "ensure you do the things right", code = {
    expect_true(TRUE)
  })
})
```

The great thing about `describe()` is that they can be *nested*: 

```{r}
describe(description = "decribe the right way to implement thing", code = {
  describe(description = "verify that you implemented the right thing", code = {
    it(description = "ensure you do the thing right", code = {
      expect_true(TRUE)
    })
  })
})
```

Each `it()` call is essentially identical to `test_that()`, but allows for a longer character string `description`. 

## What to test 

<br>

> ***Strive to test each behaviour in one and only one test. Then if that behaviour later changes you only need to update a single test*** - [What to test, R Packages, 2ed](https://r-pkgs.org/testing-design.html#what-to-test)

<br>

`pickler` attempts to address the advice above by providing a set of [Gherkin-style syntax](https://cucumber.io/docs/gherkin/reference/) functions that can be placed in the `description` argument of `decribe()` or `it()`. The three primary functions in `pickler` are `feature()`, `scenario()`, and `background()`.


The following sections walk through an example with the `split_cols()` function. You can view the source code for this function [here](https://github.com/mjfrigaard/pickler/blob/main/R/split_cols.R), but for now we'll just look it's arguments and the problem it solves: 

```{r eval=FALSE}
split_cols(
  data, 
  col, 
  pattern = "[^[:alnum:]]+", 
  col_prefix = "col")
```

Given the `input_data` data below: 

```{r , echo=FALSE}
input_data <- data.frame(id = c(1L, 2L, 3L),
                         name = c("First", 
                                  "First, Last",
                                  "First, Middle, Last"))
```

```{r}
input_data
```

`split_cols()` will generate this: 

```{r, echo=FALSE}
data.frame(id = 1:3,
  name = c("First", "First, Last", "First, Middle, Last"),
  col_1 = c("First", "First", "First"),
  col_2 = c(NA, "Last", "Middle"),
  col_3 = c(NA, NA, "Last"))
```

`split_cols()` was written because [`tidyr::separate()`](https://tidyr.tidyverse.org/reference/index.html#character-vectors) didn't have an easy solution, and I wanted to reduce the number of keystrokes/neurons/dependencies/etc to solve this problem. 

```{r, eval=FALSE}
tidyr::separate(data = input_data, 
                     col = "name", 
                     into = c("col_1", "col_2", "col_3"), 
                     remove = FALSE,
                     extra = "merge", 
                     fill = "right")
```

```{verbatim}
#>   id                name col_1  col_2 col_3
#> 1  1               First First   <NA>  <NA>
#> 2  2         First, Last First   Last  <NA>
#> 3  3 First, Middle, Last First Middle  Last
```


## Features 

Features are the units of functionality we're trying to deliver. We can adapt the language from `testthat`'s BDD functions above into a [Gherkin-style](https://cucumber.io/docs/gherkin/reference/) feature:

```gherkin
Feature: My function's feature
  As a user of this function
  I want to verify that I implemented the right thing
  So that I can ensure I did the thing right
```

Features have the following keywords: 

```gherkin
Feature: <title>
  As a <person using the code>
  I want <some action>
  So that <desired behavior>
```

A feature should start with a mental model of the problem we'd like to solve. The [Gherkin-style](https://cucumber.io/docs/gherkin/reference/) example for `split_cols()` might look something like this: 

```gherkin
Feature: Split a single column into multiple columns using a pattern
  As an analyst 
  I want to specify a column and a pattern 
  So that I can quickly generate multiple columns.
```

The goal of a feature is to clearly state the objective or functionality we're trying to achieve. The arguments in `feature()` match the Gherkin feature keywords:  

```{r, eval=FALSE}
feature(
    title = ,
    as_a = ,
    i_want = ,
    so_that = 
  )
```

This pattern is similar for all `pickler` functions, but I've converted the keywords to `snake_case` so they align with the [`tidyverse` naming conventions](https://style.tidyverse.org/syntax.html#object-names)).

`pickler`'s function arguments also serve as prompts to fill out each keyword: 

```{r eval=FALSE}
feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  )
```

You might be wondering why you'd bother passing this information into a function and not just include them in a language-agnostic fenced code block: 

````verbatim
```
Feature: Split a single column into multiple columns using a pattern
  As an analyst 
  I want to specify a column and a pattern 
  So that I can quickly generate multiple columns.
```
````

`picker`'s functions are designed to work in R Markdown *and* `testthat` tests. For example, any code blocks with a `pickler` functions can set to `eval=TRUE`, `echo=FALSE`, and `comment=""` to render nicely formatted Gherkin syntax when the file is knitted: 

```{r eval=TRUE, echo=FALSE, comment=""}
feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  )
```

## Scenarios

Examples are the cornerstone of behavior-driven development and testing. Captured as *scenarios*, these should illustrate a concrete example of how our code should behave. Scenarios are also great for communicating and documenting requirements, code, and tests with stakeholders and non-technical audiences:

<br>

> ***"Examples play a primary role in BDD, simply because they're an extremely effective way of communicating clear, precise, and unambiguous requirements. Specifications written in natural language are, as it turns out, a terribly poor way of communicating requirements, because there's so much space for ambiguity, assumptions, and misunderstandings. Examples are a great way to overcome these limitations and clarify the requirements. Examples are also a great way to explore and expand your knowledge."*** - [BDD in Action, John Ferguson Smart](https://www.manning.com/books/bdd-in-action)

<br>

Consider the `scenario()` below for `split_cols()`:

```{r eval=FALSE}
scenario(
    title = "Split column with a default pattern",
    given = "a dataframe with a specified column",
    when = "I split the column using the default pattern",
    then = "the column should be split into multiple columns"
  )
```

We could also add function arguments (in square brackets), but this is optional. 

```{r eval=FALSE}
scenario(
    title = "Split column with a default pattern",
    given = "a dataframe [data] with a specified column [col]",
    when = "I split the [col] column using the default [[^[:alnum:]]+]",
    then = "the [col] column should be split into multiple columns"
  )
```

When rendered, this scenario would look like so:

```{r eval=TRUE, echo=FALSE, comment=""}
scenario(
    title = "Split column with a default pattern",
    given = "a dataframe [data] with a specified column [col]",
    when = "I split the [col] column using the default [[^[:alnum:]]+]",
    then = "the [col] column should be split into multiple columns"
  )
```

### And

Scenarios can also include additional `and` statements after the `then` keyword to test multiple behaviors: 

```{r , eval=FALSE}
scenario(
    title = "Split column with a default pattern",
    given = "a dataframe [data] with a specified column [col]",
    when = "I split the [col] column using the default [[^[:alnum:]]+]",
    then = "the [col] column should be split into multiple columns",
    and = "the new columns should be named with the default prefix [cols_]"
  )
```

Gherkin scenarios occasionally include `And` keywords directly following a `Given` statement, but in `pickler` this information is placed in the `background()`:

## Background 

`background()` can be used to provide context for any preexisting conditions (or the state of the world) we're contending with. `background()` can also be used to reduce any repetitive information in scenarios. For example, consider the two scenarios below: 

```gherkin
  Scenario: Split column with 'default' pattern
    Given a dataframe with text data
    And a specified column "<column_name>"
    When I split the column with the default pattern
    And a column prefix "<col_prefix>"
    Then the column should be split into multiple columns
    And the resulting dataframe should have the original data with added split columns
    And the new columns should have names with the provided prefix and an index

  Scenario: Split column with 'custom' pattern
    Given a dataframe with text data
    And a specified column "<column_name>"
    When I split the column with a custom pattern "<pattern>"
    And a column prefix "<col_prefix>"
    Then the column should be split according to the custom pattern
    And the resulting dataframe should have the original data with added split columns
    And the new columns should have names with the provided prefix and an index
```

These scenarios have a lot of duplicated lines--specifically, these three: 

```gherkin
Given a dataframe with text data
And a specified column "<column_name>"
And a column prefix "<col_prefix>"
```

We can place these in a `background()` to remove repetition and 'set the stage' for any future scenarios:

```{r eval=FALSE}
background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
```

The nice thing about storing additional context in `background()` is that you'll only have to update this information in a single location.

When rendered, the `background()` looks like this:

```{r eval=TRUE, echo=FALSE, comment=""}
background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
```

## Tests

BDD tests are written to communicate the context, intended behavior, and a series of use cases for how our code should be used. `pickler` functions help facilitate this because they can be placed in nested calls to `describe()` and `it()`. 

Below is an example `feature()`, `background()`, and `scenario()` for `split_cols()`:

```{r}
describe(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ), code = {
    
    describe(
      background(
        title = "Input dataframe with text data", 
        given = "a dataframe [data] with text columns", 
        and = list(
                   "a column prefix [cols_]", 
                   "a specified column [col]")), code = {
    
    it(
      scenario(
        title = "Split column with a default pattern",
        given = "a dataframe [data] with a specified column [col]",
        when = "I split the [col] column using the default [[^[:alnum:]]+]",
        then = "the [col] column should be split into multiple columns"
      ), code = {
      
      expect_true(TRUE)
      
    })
  
  })
    
})
```

BDD tests take longer to write, but the tradeoff is obvious: any losses in brevity are gained in clarity. The emphasis on using plain language makes it easier for stakeholders and collaborators to understand how our code works (and how it's being tested).

`pickler` functions can also be placed in `desc` argument of a the `test_that()` test, but sometimes a multiline descriptions [can cause errors](https://github.com/r-lib/testthat/issues/1900#issue-1995181567).: 

```{r}
test_that(
  feature(
    title = "My function's feature",
    as_a = "user of this function",
    i_want = "to verify that I implemented the right thing",
    so_that = "to ensure I did the thing right"
  ),
  code = {
    expect_true(TRUE)
  }
)
```

## Bundling 

Ideally, `pickler`'s functions are developed independently and then wrapped in a `bundle()` before passing to the `testthat` BDD functions: 

```{r , eval=FALSE}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
)
```

Bundling makes it easier to combine features and backgrounds into a single call:

```{r , echo=FALSE, eval=TRUE, comment=""}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
)
```

`bundle()` also allows us to pass a `feature()` and `background()` to `describe()`, then test a particular `scenario()` with `it()`:

```{r}
describe(
  # bundle feature and background
  bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
), code = {
    it(
    # test scenario 1
     scenario(
        title = "Split column with a default pattern",
        given = "a dataframe [data] with a specified column [col]",
        when = "I split the [col] column using the default [[^[:alnum:]]+]",
        then = "the [col] column should be split into multiple columns"
      ), code = {
      
      expect_true(TRUE)
      
    })
    it(
    # test scenario 2
     scenario(
        title = "Split column with a default pattern",
        given = "a dataframe [data] with a specified column [col]",
        when = "I split the [col] column using the default [[^[:alnum:]]+]",
        then = "the new columns should be named with the default prefix [cols_]"
      ), code = {
      
      expect_true(TRUE)
      
    })
  })
```


## Tables

It's also possible to provide a table in a scenario or feature. Let's assume we have a small example of data that illustrates an example behavior for our `split_cols()` function (stored in `input`):

```{r , echo=FALSE}
input <- data.frame(
  value = c(1L, 2L, 3L),
  name = c(
    "John", 
    "John, Jacob",
    "John, Jacob, Jingleheimer"
  )
)
```

```{r, comment="#>"}
input
```

We can convert `input` to a Gherkin-style table using `with_table()`:

```{r eval=FALSE}
with_table(input)
```

```{r echo=FALSE, eval=TRUE, comment=""}
with_table(input)
```

`with_table()` creates a slightly modified `knitr::kable(format = "pipe", align = "l")` table we can include in a `bundle()` with the `background` and `feature`:

```{r , eval=FALSE}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [name]", 
               "a specified column [name]")),
  
  input = c("
           |value |name                      |
           |------|--------------------------|
           |1     |John                      |
           |2     |John, Jacob               |
           |3     |John, Jacob, Jingleheimer |
           ")
)
```

This makes it easier to view them in RMarkdown or Quarto files (with `echo` set to `FALSE` and `comment` set to `""`)

```{r , echo=FALSE, comment=""}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [default]", 
               "a specified column [name]")),
  
  input = c("
           |value |name                      |
           |------|--------------------------|
           |1     |John                      |
           |2     |John, Jacob               |
           |3     |John, Jacob, Jingleheimer |
           ")
)
```

If we know we know what we the `output` table should look like, we can include it with the `scenario`: 

```{r , echo = FALSE}
output <- 
data.frame(value = 1:3, name = c("John", "John, Jacob", "John, Jacob, Jingleheimer"
), col_1 = c("John", "John", "John"), col_2 = c(NA, "Jacob", 
"Jacob"), col_3 = c(NA, NA, "Jingleheimer"))
```

```{r , eval=FALSE, echo=TRUE}
with_table(output)
```

```{r , eval=TRUE, echo=FALSE, comment=""}
with_table(output)
```


```{r , eval=FALSE}
bundle(
scenario(
    title = "Split column with a default pattern",
    given = "a dataframe [data] with a specified column [name]",
    when = "I split the [name] column using the default [[^[:alnum:]]+]",
    then = "Then the [name] column should be split into multiple columns",
    and = "the new columns should be named with the default prefix [cols_]"
  ),
output = c(
          "
          |value |name                      |col_1 |col_2 |col_3        |
          |------|--------------------------|------|------|-------------|
          |1     |John                      |John  |NA    |NA           |
          |2     |John, Jacob               |John  |Jacob |NA           |
          |3     |John, Jacob, Jingleheimer |John  |Jacob |Jingleheimer |
          ")
)
```

Now our `background`, `feature`, `scenario` (and tables) are in the Gherkin-style format that can be placed in a R Markdown file and nicely rendered for stakeholders or non-technical audiences: 

```{r , eval=TRUE, echo=FALSE, comment=""}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [default]", 
               "a specified column [name]")),
  
  input = c("
           |value |name                      |
           |------|--------------------------|
           |1     |John                      |
           |2     |John, Jacob               |
           |3     |John, Jacob, Jingleheimer |
           "),
  scenario(
      title = "Split column with a default pattern",
      given = "a dataframe [data] with a specified column [name]",
      when = "I split the [name] column using the default [[^[:alnum:]]+]",
      then = "the [name] column should be split into multiple columns",
      and = "the new columns should be named with the default prefix [cols_]"
  ),
output = c(
          "
          |value |name                      |col_1 |col_2 |col_3        |
          |------|--------------------------|------|------|-------------|
          |1     |John                      |John  |NA    |NA           |
          |2     |John, Jacob               |John  |Jacob |NA           |
          |3     |John, Jacob, Jingleheimer |John  |Jacob |Jingleheimer |
          ")
)
```

And our test file contains a fully scoped test with everything documented:

```{r}
describe(
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [default]", 
               "a specified column [name]")),
  
  input = c("
           |value |name                      |
           |------|--------------------------|
           |1     |John                      |
           |2     |John, Jacob               |
           |3     |John, Jacob, Jingleheimer |
           ")), code = {
    
it(
 bundle(
  scenario(
      title = "Split column with a default pattern",
      given = "a dataframe [data] with a specified column [name]",
      when = "I split the [name] column using the default [[^[:alnum:]]+]",
      then = "the [name] column should be split into multiple columns",
      and = "the new columns should be named with the default prefix [cols_]"
  ),
output = c("
          |value |name                      |col_1 |col_2 |col_3        |
          |------|--------------------------|------|------|-------------|
          |1     |John                      |John  |NA    |NA           |
          |2     |John, Jacob               |John  |Jacob |NA           |
          |3     |John, Jacob, Jingleheimer |John  |Jacob |Jingleheimer |
        ")
  ), code = {
  
  # create observed output
  observed <- split_cols(data = input, col = 'name')
  # compare against output
  expect_equal(object = observed, expected = output)
  
})

})
```

And when it's placed in the [test file](https://github.com/mjfrigaard/pickler/blob/main/tests/testthat/test-split_cols.R), it passes!

```{r , eval=FALSE}
devtools:::test_active_file()
```

```{verbatim}
[ FAIL 0 | WARN 0 | SKIP 0 | PASS 1 ]
```

