---
title: "Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(testthat)
```


```{r setup}
library(pickler)
```

Make sure you have the latest version of `testthat` installed.

```{r testthat, eval=FALSE, message=FALSE, warning=FALSE, results='hide'}
if (!require(pak)) {
  install.packages("pak",
    repos = "http://cran.us.r-project.org"
  )
}
pak::pkg_install("r-lib/testthat", upgrade = TRUE, ask = FALSE)
library(testthat)
```


## Tables

It's also possible to provide a table in a scenario or feature. Let's assume we have a small `input` that illustrates an example behavior for our `pivot_string_long()` function:

```{r , echo=FALSE}
input <- c(input = "one-two-three")
```

```{r, comment="#>"}
input
```

We can include `input` in our bundle with the `feature()` and `scenario()`:

```{r , eval=FALSE}
bundle(
  feature(
    title = "Pivot String Long",
    as_a = "user of pivot_string_long()",
    i_want = "to specify a text column and a pattern",
    so_that = "peform computations on the unique items."
  ),
  background(
    title = "Input text data", 
    given = "a string or vector [string] with text columns", 
    and = "a specified pattern [sep]"),
  
  input = c("one-two-three")
)
```

```{r , echo=FALSE, comment=""}
bundle(
  feature(
    title = "Pivot String Long",
    as_a = "user of pivot_string_long()",
    i_want = "to specify a text column and a pattern",
    so_that = "peform computations on the unique items."
  ),
  background(
    title = "Input text data", 
    given = "a string or vector [string] with text columns", 
    and = "a specified pattern [sep]"),
  
  input = c("one-two-three")
)
```

If we know we know what we the `output` table should look like, we can include it with the `scenario`: We can convert `output` to a Gherkin-style table using `with_table()`:

```{r , echo = FALSE}
output <- data.frame(
    unique_items = c("one", "two", "three"),
    string = c("one-two-three", NA, NA))
```

```{r , eval=FALSE, echo=TRUE}
with_table(output)
```

```{r , eval=TRUE, echo=FALSE, comment=""}
with_table(output)
```


`with_table()` creates a slightly modified `knitr::kable(format = "pipe", align = "l")` table we can include in a `bundle()` with the `background` and `feature`:

```{r , eval=FALSE}
bundle(
scenario(
    title = "Split a single string with default separator",
    given = "a character [string] 'one-two-three'",
    when = "I pass the [string] to pivot_string_long()",
    then = "Then [unique_items] column should contain rows: 'one, two, three'",
    and = "[string] column should contain the original 'one-two-three'"
  ),
output = c("
      |unique_items |string        |
      |-------------|--------------|
      |one          |one-two-three |
      |two          |NA            |
      |three        |NA            |
          ")
)
```

This makes it easier to view them in RMarkdown or Quarto files (with `echo` set to `FALSE` and `comment` set to `""`)

Now our `background`, `feature`, `scenario` (and tables) are in the Gherkin-style format that can be placed in a R Markdown file and nicely rendered for stakeholders or non-technical audiences: 

```{r , eval=TRUE, echo=FALSE, comment=""}
bundle(
  feature(
    title = "Pivot String Long",
    as_a = "user of pivot_string_long()",
    i_want = "to specify a text column and a pattern",
    so_that = "peform computations on the unique items."
  ),
  background(
    title = "Input text data", 
    given = "a string or vector [string] with text columns", 
    and = "a specified pattern [sep]"),
  
  input = c("one-two-three"),
  
  scenario(
    title = "Split a single string with default separator",
    given = "a character [string] 'one-two-three'",
    when = "I pass the [string] to pivot_string_long()",
    then = "Then [unique_items] column should contain rows: 'one, two, three'",
    and = "[string] column should contain the original 'one-two-three'"
  ),
  output = c("
      |unique_items |string        |
      |-------------|--------------|
      |one          |one-two-three |
      |two          |NA            |
      |three        |NA            |")
)
```

Now our [test file](https://github.com/mjfrigaard/pickler/blob/main/tests/testthat/test-pivot_string_long.R) contains a fully scoped test with everything documented:

```{r}
describe(
bundle(
  feature(
    title = "Pivot String Long",
    as_a = "user of pivot_string_long()",
    i_want = "to specify a text column and a pattern",
    so_that = "peform computations on the unique items."
  ),
  background(
    title = "Input text data",
    given = "a string or vector [string] with text columns",
    and = "a specified pattern [sep]"),

  input = c("one-two-three")), code = {

it(
 bundle(
  scenario(
    title = "Split a single string with default separator",
    given = "a character [string] 'one-two-three'",
    when = "I pass the [string] to pivot_string_long()",
    then = "Then [unique_items] column should contain rows: 'one, two, three'",
    and = "[string] column should contain the original 'one-two-three'"
  ),
  output = c("
      |unique_items |string        |
      |-------------|--------------|
      |one          |one-two-three |
      |two          |NA            |
      |three        |NA            |")
   ), code = {
        
  input <- c("one-two-three")
  # create observed output
  observed <- pivot_string_long(input)
  
  # compare against output
  output <- data.frame(
    unique_items = c("one", "two", "three"),
    string = c("one-two-three", NA, NA))
  
  expect_equal(object = observed, expected = output)

})

})
```

And... 

```{r , eval=FALSE}
devtools:::test_active_file()
```

it passes!

```{verbatim}
[ FAIL 0 | WARN 0 | SKIP 0 | PASS 1 ]
```

