---
title: "Workflow"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
# Make sure you have the latest version of `estthat and sloop installed.
# if (!require(sloop)) {
#   install.packages("sloop",
#     repos = "http://cran.us.r-project.org"
#   )
# }
# pak::pkg_install("r-lib/sloop", upgrade = TRUE, ask = FALSE)
# 
# if (!require(pak)) {
#   install.packages("pak",
#     repos = "http://cran.us.r-project.org"
#   )
# }
# pak::pkg_install("r-lib/testthat", upgrade = TRUE, ask = FALSE)
```

```{r pkgs}
library(pickler)
library(testthat)
```


## Test development workflow

In standard R packages, new functionality is added to `R/` with `usethis::use_r()`, followed by a corresponding test file in `tests/testthat/` with `usethis::use_test()`:

-   [x] `use_r()`

-   [x] `use_test()`

While developing, these files can be accessed using the functions above, and tested interactively with `devtools::test_active_file()` and `devtools::test_coverage_active_file()`

-   [x] `test_active_file()`

-   [x] `test_coverage_active_file()`

As the code in `R/` grows, the test suite might include additional features like fixtures, helpers, mocks, or a `setup.R` file. When a package has been sufficiently developed, the tests are run with `devtools::test()`, test coverage is reported with `devtools::test_coverage()`, and the (I'm assuming the package is checked periodically with `devtools::check()`, too):

-   [x] `test()`

-   [x] `test_coverage()`

Additional package can be used to include test coverage reports ([`covrpage`](https://yonicd.github.io/covrpage/)), traceability matrices ([`covtracer`](https://genentech.github.io/covtracer/)), etc., to your package documentation.

## Test documentation

Throughout development, the functional requirements for your package are documented in `R/` (with `roxygen2`), and long-form documentation is placed in vignettes.[^r-pkgs-vignettes] Vignettes can (and should) be used to show users how your package works, especially if your package is going to be shared with the world (i.e., via a [`pkgdown` site](https://pkgdown.r-lib.org/)). However, it's not often developers use vignettes to store the information that's essential to the development *process* (i.e., proposed features, specifications, user stories, etc.) 

[^r-pkgs-vignettes]: The [vignettes](https://r-pkgs.org/vignettes.html#advice-on-writing-vignettes) chapter in R Packages, 2ed is a wonderful guide on writing that extends far beyond R packages.


## Rmd + BDD = `pickler` 

Including relevant context in `describe()` and `it()` is nice for developers, but I've found stakeholders and potential users rarely look the the `tests/` folder for information on a package's features (or how it's functions are supposed to behave).

`pickler` was written to make the information in the `tests/` folder more accessible to a broader audience. The [Gherkin-style](https://cucumber.io/docs/gherkin/reference/) functions available in `pickler` are designed to work in R Markdown (i.e. as a vignette) *and* run in a test file. It's been pointed out elsewhere[^2] that vignettes have a few characteristics that make them a great place for development:

<br>

> ***"an R package implies writing documentation: one of the main goals of the Vignettes, in an R package, is to document how to use the package. And the good news is that when checking a package, i.e. when running `check()` from `devtools` ([Wickham, Hester, and Chang 2020](https://devtools.r-lib.org/)) or `R CMD check`, the Vignettes are going to be built, and the process will fail if at least one of the Vignettes fails to render. That way, you can use the documentation of the back-end as an extra tool for doing unit testing!"***

<br>

[^2]: This point in made in [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/building-ispum-app.html#rmd-vignettes-and-documentation-first), and although the text is geared towards developing a Shiny app-package, the quote is also applicable to non-Shiny R packages. 


```{r , eval=FALSE, include=FALSE, comment=""}
feature(
    title = "Text-based Logo Generation",
    as_a = "As a user who calls the text_logo() function",
    i_want = "to generate a text-based logo",
    so_that = "I can quickly insert the pickler logo"
    )
```
