---
title: "Bundling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bundling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(testthat)
```

```{r setup}
library(pickler)
```

Make sure you have the latest version of `testthat` installed.

```{r testthat, eval=FALSE, message=FALSE, warning=FALSE, results='hide'}
if (!require(pak)) {
  install.packages("pak",
    repos = "http://cran.us.r-project.org"
  )
}
pak::pkg_install("r-lib/testthat", upgrade = TRUE, ask = FALSE)
library(testthat)
```


## Bundling 

Ideally, `pickler`'s functions are developed independently and then wrapped in a `bundle()` before passing to the `testthat` BDD functions. Consider the `feature()` and `scenario()` below for the `split_cols()` function:

```{r , eval=FALSE}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
)
```

Bundling makes it easier to combine features and backgrounds into a single call:

```{r , echo=FALSE, eval=TRUE, comment=""}
bundle(
  feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
  background(
    title = "Input dataframe with text data", 
    given = "a dataframe [data] with text columns", 
    and = list(
               "a column prefix [cols_]", 
               "a specified column [col]"))
)
```

`bundle()` also allows us to pass a `feature()` and `background()` to `describe()`, then test a particular `scenario()` with `it()`:

```{r}
describe(
  bundle(
    feature(
    title = "Split a single column into multiple columns using a pattern",
    as_a = "user of split_cols()",
    i_want = "to specify a column and a pattern",
    so_that = "I can quickly generate multiple columns."
  ),
    background(
    title = "Input dataframe with text data",
    given = "a dataframe [data] with text columns",
    and = list(
               "a column prefix [default]",
               "a specified column [name]")),

  input = c("
           |value |name                      |
           |------|--------------------------|
           |1     |John                      |
           |2     |John, Jacob               |
           |3     |John, Jacob, Jingleheimer |
           ")), code = {

  it(
  bundle(
    scenario(
      title = "Split column with a default pattern",
      given = "a dataframe [data] with a specified column [name]",
      when = "I split the [name] column using the default [[^[:alnum:]]+]",
      then = "the [name] column should be split into multiple columns",
      and = "the new columns should be named with the default prefix [cols_]"
  ),
  output = c("
          |value |name                      |col_1 |col_2 |col_3        |
          |------|--------------------------|------|------|-------------|
          |1     |John                      |John  |NA    |NA           |
          |2     |John, Jacob               |John  |Jacob |NA           |
          |3     |John, Jacob, Jingleheimer |John  |Jacob |Jingleheimer |
        ")
  ), code = {
  input <- data.frame(
              value = c(1L, 2L, 3L),
            name = c(
              "John",
              "John, Jacob",
              "John, Jacob, Jingleheimer"
            )
    )
  # create observed output
  observed <- split_cols(data = input, col = 'name')
  # create expected output 
  output <- data.frame(
    value = 1:3,
    name = c("John", "John, Jacob", "John, Jacob, Jingleheimer"),
    col_1 = c("John", "John", "John"),
    col_2 = c(NA, "Jacob", "Jacob"),
    col_3 = c(NA, NA, "Jingleheimer"))
  # compare against output
  expect_equal(object = observed, expected = observed)
  })
})
```

